A High Efficiency MAC Protocol for WLANs: Providing Fairness in Dense Scenarios

### 2015-11-09 responseLetter.tex

		In case $\sigma_{e}$ (empty slot duration) and $\sigma_{s}$ (duration of slots containing a transmission) are of similar size, the reviewer is right. For instance, we roughly have that the aggregate throughput once the collision-free operation has been reached is $S(2)= 2L/(6\sigma_{e} + 2\sigma_{s})$ and $S(3)= 3L/(5\sigma_{e} + 3\sigma_{s})$ for 2 and 3 contenders respectively. Given that $\sigma_{s} = \sigma_{e}$, we get that $S(3)/S(2) = 1.5$, as indicated. However, since $\sigma_{s} \approx 1000\sigma_{e}$, the ratio $S(3)/S(2)$ is bigger than 1, but relatively small.

		In later figures, this also explains why once in collision free, with few STAs we almost reach the maximum throughput, and the increase when more STAs are added is small. See for example, the throughput of CSMA/ECA in Figure 2.

@@

		Yes. When the MAC queue is emptied, CSMA/ECA nodes reset their backoff stage and retransmission counters to zero ($k$ and $r$, respectively). This means that in non-saturated traffic, Schedule Reset is of little use. Figure~\ref{fig:queueSize} shows the average number of packets in the MAC queue for CSMA/CA and CSMA/ECA.

		\begin{figure}[tb]
		\centering
			\includegraphics[width=0.45\linewidth,angle=-90]{figures/unsaturated/queueSize/queueSize-multiplot-TON.eps}
			\caption{Average MAC queue size in non-saturation}
			\label{fig:queueSize}
		\end{figure}

	\subsection{Regarding ``Minor"}
		\begin{itemize}
			\item {\bfseries Another idea comes to the reviewers mind. If k and CW\_min can be adjusted smartly, for sparse scenario, is the following schedule possible?
    					STA 1:  1   t   1   t;
     					STA 2:  t   1   t   1; 
			i.e., k == 0 and CW\_min == 0. (The reviewer might be wrong.)}
		\end{itemize}

		Having fast and accurate information about the number of contenders in WLANs is a challenging task. In fact, failure to accurately estimate this number is one of the drawbacks of protocols like L-MAC and ZC-MAC.

		To answer the question: yes, it is possible. Nevertheless, as CSMA/ECA is though to be an evolution of CSMA/CA, we do not use those CW$_{\min}$ values so we comply with the standard.

		Furthermore, to make CSMA/ECA compatible with CSMA/CA for coexistence scenarios, the value of the deterministic backoff is roughly the expected value of CSMA/CA's random backoff at backoff stage $k$. That is, $B_{\text{d}}\coloneqq\lceil{E[0,CW(k)-1]}\rceil$ (as described in Section III-H in the paper).

@@

		\begin{itemize}
			\item {\bfseries The reviewer is interested in the k values in high contention scenarios. If not too difficult, please make figures for k, especially when CSMA/ECA\_{Hys+FS} co-exist with CSMA/CA. Does CSMA/CA nodes cause collision and increases k? If not, why not?}
		\end{itemize}

		Now that Schedule Reset is introduced we acknowledge that collisions with CSMA/CA nodes, channel errors, slots drifts and addition/withdrawals of contenders will increase CSMA/ECA$_{\text{Hys+FS}}$ nodes's schedule length to its maximum value, that is, the average backoff stage will rapidly grow to the maximum value, $k=m$. This produces an increase in the time between successful transmissions, which the node attempts to reduce with Schedule Reset.
